### 2020-11-24 DissertationProposal/DissertationProposal.Rmd
Totally redid abm descriotion

--- 
title: Cultural Evolution of Conservation Behavior at the Intersection of Working
  Landscapes and Protected Areas
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: yes
  pdf_document:
    toc: yes
    extra_dependencies: "subfig"
  bookdown::html_document2:
    toc: yes
fontsize: 12pt
link-citations: yes
csl: AmJBot.csl
subtitle: Dissertation Proposal for Committee Review
bibliography: [References/Dissertationproposal.bib,References/NonZoteroRefs.bib]
---
```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')
```

\newpage
# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common, however, for protected areas to fail to reach their stated conservation goals. The most chronic cause of conservation failure is a lack of local support and adoption of mandates. Conversely, historic protected area successes are also attributed to widespread adoption of conservation behaviors. Research examining how and why protected areas are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin local adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

1. Develop a dynamic, theoretical model to generate predictions regarding if and how group-level adoption of protected areas can emerge and persist via competition between groups.

2. Empirically test theoretical predictions of protected area adoption as a function of group competition and perceived costs and benefits to local community members.

3. Using the ongoing spread of an improved efficiency cookstove design, examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.

Globally, we have expended a tremendous amount of resources to identify conservation goals which will deliver the greatest benefits to biodiversity and ecosystem function. For these benefits to come to fruition however, conservation initiatives must be adopted at a rate which exceeds ecological degradation. This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.  
\newpage

# Introduction
## Phenomenon
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources [@network2019global]. Today, on average, we degenerate these resources at a rate 56% faster than regeneration [@network2020global]. This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993]. Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity [@regev_biological_1998].

The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity [@haines2010links; @redford1995human]. As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target [@parrish_are_2003]. Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet]. These area-based conservation initiatives are from hereon referred to as protected areas.

```{r paNumber, echo=FALSE, fig.cap="Number of newly established no-take protected areas reported by the IUCN 2020", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/Number.png")
```

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010]. While high-profile protected area successes are salient in the conservation narrative, protected areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is somewhat speculative as unsuccessful protected areas go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004]. Regardless of the high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].

Community norms around protected area adherence are characterized as individually costly, yet beneficial at the group level. Substantial variation exists in the degree to which communities adhere to protected area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@rogers_diffusion_2010; @simon_win-win_2012]. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017]. This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution]. What is less understood however, is the emergence and initial adoption of conservation behaviors.

Research on the diffusion of conservation behaviors suffers from a lack of generalizability. Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors. Outside of conservation science, the field of evolutionary anthropology has devoted considerable effort to understanding how and why individuals adopt costly cooperative behaviors. Of particular relevance to understanding group adherence to protected area regulations, the theory of cultural group selection offers a mechanistic explanation for the emergence and spread of individually costly group norms.

## Primer on cultural group selection
The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations [@boyd_transmission_2010]. Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not. As an individual’s behavior is flexible, individuals may migrate from less to more successful groups or directly copy the behaviors they observe in visibly successful cooperative groups [@henrich_cultural_2004]. This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection. In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].

Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008]. Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically. Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.

## Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004]. While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005]. Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

## Fidelity of behavioral transmission
The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission. In situations where behaviors do emerge and spread within populations, behaviors are transmitted along a chain of individuals, where they may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011]. This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves. Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator [@king2007use; @muthukrishna_when_2016]. These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.

## Dissertation objectives
The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge and spread efficiently within populations?

* Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation. Formally test the intuition that between group imitation can lead to the spread of community wide protected area adoption whereas solitary groups may fail to cooperate under identical environmental constraints such as resource degradation, regeneration, and mobility.

* Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania. Probabilities are based on the salience of group cultural identities and the costs and benefits of CoFMAs to individuals and groups. Compare theoretical predictions to the presence or absence of CoFMAs in reality.

* Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior. This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.

## Study System
All empirical research for this dissertation will take place on Pemba Island, Tanzania. In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar]. This population is divided into wards of 2,000 to 6,000 individuals. The majority of Pemban wards are described as forest-dependent. The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, establish forest protected areas and exclude outside individuals [@benjaminsen2014between]. As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].

\newpage

# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
## Background
Cultural group selection theorizes that individuals will adopt cooperative behavior if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985]. Under this framework, the presence of multiple competing groups drives community cooperative norms to emerge under a much broader set of conditions compared to solitary groups [@richerson_cultural_2016]. Protected area conservation is a cooperative phenomenon where individuals sacrifice some degree of their own harvest potential in order to benefit the group as a whole. Therefore, insights from cultural group selection may help conservation scientists better understand and predict community adherence to protected area regulations [@brooks_applying_2018].

While cultural group selection theory offers mechanistic insights for the emergence of cooperative norms in groups broadly, its applicability to conservation specifically is unknown. Conservation has many aspects which make it unique compared to cooperation in the abstract. In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making. Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as mobility and regeneration rate. These specifics make conservation a fundamentally different system from general cooperation, thus we must extend cultural group selection theory considering conservation explicitly.

This project aims to logically test the applicability of cultural group selection to conservation. Specifically, this research will address how behavior-environment feedbacks and resource mobility and regeneration rates affect the emergence and spread of protected area adoption. In this chapter I will develop a dynamic, agent-based simulation of individuals harvesting resources from a landscape under a variety of environmental constraints. I will then use this simulation to test under what circumstances adherence to protected area regulations emerges as a group norm within solitary and multiple, competing groups. This model will generate qualitative predictions about system outcomes given a very simple set of assumptions.

### Question:
Under what environmental contexts might the mechanisms of cultural group selection influence individual decision making to generate lasting group norms around the adherence to protected area regulations?

### Predictions:
* P1: Groups will adopt protected area regulations in both the solitary and multiple competing groups scenarios under specific environmental contexts.

* P2: In the presence of multiple competing groups, the environmental contexts which generate group adoption of protected area regulations will be much more broad compared to solitary groups.

## Research Design

An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual [@railsback_agent-based_2012; @wilensky_introduction_2015]. This design mirrors our real-world understanding of the important system components.

### Model Description

*Current Single Group Model - The Most Simple Case(s)*

In the most basic form, the current agent-based model demonstrates repeated subtractions from a multiplicatively regenerating quantity of resources. In this context, there are no resource protections and therefore agents cannot modify their harvest strategies. We can write a simple analytical model to better denote these dynamics and determine interesting and appropriate starting conditions for more complex permutations of the agent-based model, as well as validate our intuitions as to how the model should behave. 
Given that the resources regenerate multiplicatively, any starting conditions which cause the resource pool to diminish after the first time step will further deplete the ever-shrinking resource pool with each additional time step. Specifically, any max harvest value which when multiplied by the number of agents is greater than the number of resources that regenerate in the first time step, will cause the resource pool to continually diminish over time and eventually reach zero. We can define the sustainable harvest threshold (${H_{sust}}$) for any given number of agents (${N}$), starting resources (${W_{r}}$), and regeneration rate (${R}$) as:

$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)}{N} $$

Figure \@ref(fig:FirstABM) shows the model behavior when the maximum harvest value is at the calculated sustainable harvest value for this scenario. We can see that the amount of resources harvested in each time step is below the amount regenerated, therefore we see no change in the available and harvested resources over time.

```{r FirstABM, out.width = "100%",echo=FALSE, fig.align='center', dpi=72, fig.pos= "H",fig.cap="Resource quantity and mean harvest values over time for one run of the agent-based model at a maximum harvest value exactly at the calculated value for sustainable harvest. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20."}
knitr::include_graphics("Figures/Ch1/abmP1a.png")
```

In contrast to the figure above, figure \@ref(fig:FirstABMb) shows the model behavior when the maximum harvest value is just one over the calculated sustainable harvest value. In this scenario, we can see that the resource pool becomes more depleted over time and the quantity of resources regenerated at each time step is reduced proportionally.

```{r FirstABMb, out.width = "100%",echo=FALSE, fig.cap="Resource quantity and mean harvest vales over time for one run of the agent-based model at a maximum harvest valuue one above the calculated value for sustainable harvest. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmP1.png")
```

We can extend this model to explore how the sink/source dynamics between protected areas and adjacent working landscapes may provision more renewable resources than working landscapes alone. In this formation, we still have one homogenous group of agents who harvest regenerating resources from a landscape. This landscape now has some specified percentage sectioned off and inaccessible to the agents — the protected area. At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount. The outcome of this harvest attempt for each agent is the result of one random draw from a Binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e. that individual’s share of the available resources) and the probability of success is equal to the percent of an agent’s harvest time spent in the working landscape; in this scenario we are forcing obligatory cooperation so that percentage is always 100%. This will not always be the case as we build complexity in the model.

After the agents harvest resources from the landscape, the resources in the working and protected areas will reproduce by the specified regeneration factor. If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e. not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a Binomial distribution. Resources which move to the adjacent area but exceed the carrying capacity will also die. If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the protected area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).

$$P_{mob} = (R*P_{r})-P_{r}$$

$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{P_{mob},W_{cc}-W_{r}\}}{N} $$

Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area. We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (fig \@ref(fig:abm2)). In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e. all protected). Above the maximum sustainable harvest value, the size of the protected area becomes greatly important for determining average harvest payoff to agents. In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the protected area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.

```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and protected area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp2.png")
```

*Current Single Group Model - Incorporating Uncertainty*

Up to this point, we have assumed that excess resources produced in the protected area will move to the working landscape, up until a specified carrying capacity, with 100% probability. Key insights from this model are gained from experimenting with different probabilities of resource mobility. To frame this in the context of real world protected areas, we can assume that insights from model runs with high probabilities of resource mobility may be applied to wide ranging organisms such as large aquatic animals, while model runs with low probabilities of resource mobility may be applied to relatively low vagility organisms such short or large-seeded trees [@thomson2011seed].

The probability of resource mobility is operationalized in the model by using random draws from a Binomial distribution. For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility. With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between protected area size and resource mobility given obligatory cooperation.

```{r abm3, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model. Outcomes are the result of many combinations of resource mobilities and protected area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp3.png")
```

*Current Single Group Model - Cooperation, Defection, & Social Learning*

To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around protected area regulations. During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting). The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (fig \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.

```{r betastart, out.width = "75%",echo=FALSE, fig.align='center', fig.pos= "H", fig.cap="One thousand random draws from a beta distribution with the shape 1 and 2 parameters set at 1 and 2 respectively. This distribution is used to assign the harvest effort spent in the protected area for defecting agents during model initialization."}
hist(rbeta(1000,1,2),main=NULL)
```

At each time step, all agents first harvest resources from the working landscape. The number of resources agents receive from the working landscape is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the working landscape divided by the number of agents and the probability of success is equal to the percentage of effort that agent spends in the working landscape. Agents only harvest resources up to the specified maximum harvest value.

After harvesting from the working landscape, agents who defect to some degree and did not already reach the maximum harvest value, spend the remainder of their harvest effort in the protected area. Following the same logic as harvest from the working landscape, the number of resources agents receive from the protected area is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the protected area divided by the number of agents who defect in that time step and the probability of success is equal to the percentage of effort that agent spends in the protected area.

Agents modify their behavior (the percentage of time they spend in each area) at each time step according to one of two learning strategies. If agents are following a success-biased learning strategy, they are paired with another agent at random at each time step and compare strategies. If one agent performs worse than another in the previous time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them. If agents are following a conformist-biased strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group. The learning strategy of the overall population is currently set to either success-biased or conformist-biased learning for all agents, but in future versions of the model, it will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents. The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars)).

```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/ABMConceptDiagramFinal.png")
```

With all model parameters kept constant as in the output shown in (fig \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning. The average final time step outcomes from five runs of this model can be seen in (fig \@ref(fig:abm4)).

```{r abm4, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model. Outcomes are the result of many combinations of resource mobilities and protected area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25. Five percent of agents assigned to defect at initialization. Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp4.png")
```
Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes. To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (fig \@ref(fig:abm3)) and observe the model behavior at each time step.

```{r abm5, out.width = "100%",echo=FALSE, fig.cap="Five runs of the agent-based model. Thin black lines show each run. Thick black lines show the average values across runs. Starting conditions include 100 individuals harvesting from a landscape of 10000 resources. Eighty percent of resources are inside the protected area and the probability of resource mobility is 100 percent. The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25. Five percent of agents assigned to defect at initialization. Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/ABM_Output.png")
```

Model dynamics show that over time, agents begin to harvest from the protected area as they copy more successful individuals. This behavior depletes the resource pool in the protected area and lowers the number of resources which regenerate and agents can harvest in the subsequent time step. Note that full defection is defined as agents spending an equal proportion of their effort in the protected area as the proportion protected (i.e. they treat the entire landscape equally).

The results displayed above are for one homogenous group of agents harvesting resources from a common pool of resources. This model is useful for exploring the sink/source utility of protected areas to ensure long-term resource provisioning, especially in the contexts of social learning. To further explore this model, I have created an interactive application which performs one run of the model and can be found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/). Comparing these model runs to simple analytical models and our intuition about how the model should behave also allows us to ensure that the model is functioning properly. To address the research question of interest however, we will need to update the model to include multiple interacting groups.

```{r tabAbmVars, echo=F, fig.align='center'}
abmVars <- data.frame(Parameters = c("Individuals","Total Carrying Capacity", "Percent Carrying Capacity", 
                                    "Percent Protected", "Starting Cooperators", "Time Steps", "Max Harvest",
                                     "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
                                     "Percent Cooperation", "Average Payoff", "Protected Area Degradation",
                                     "Working Landscape Degradation"), 
                      Type = c("Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Key Independent Variable","Key Independent Variable",
                               "Key Independent Variable","Key Independent Variable",
                               "Outcome","Outcome","Outcome","Outcome"), 
                      Description = linebreak(c("Number of agents in simulation",
                                      "Maximum number of resources allowed on landscape",
                                      "Percent of carrying capacity at initialization",
                                      "Percent of landscape in protected area",
                                      "Starting percent of agents fully cooperating",
                                      "Number of time steps simulation will run for",
                                      "Maximum number of resources agents will harvest during one time step",
                                      "The percent of available resources which will regenerate at each time step in both protected and working landscapes",
                                      "The probability that any available resource will move from the  protected to working landscape or vice versa",
                                      "Solitary or multiple competing groups",
                                      "Assigns a percent of the population to success-biased imitation and the remaining population to conformist-biased imitation",
                                      "Percent of total population harvest time spent in the working landscape","Mean harvest for all agents at each time step",
                                      "Working landscape percent of total carrying capacity",
                                      "Protected area percent of total carrying capacity")))
kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F) %>% 
  kable_classic_2(full_width=F)%>% row_spec(0,bold=TRUE)%>%
    column_spec(1, width = "8em")%>%
    column_spec(2, width = "8em")%>%
  column_spec(3, width = "20em")

```

*Proposed Multi-Group Model*

In the multiple interacting groups scenario, two groups with separate resource pools will each follow the processes described above. However, at the end of each time step the groups will compare the average harvest rates. We assume that the underperforming group can observe the average harvest strategy (time spent in the protected or working areas) of the better performing group. A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).

Overall, the goal of this model is to test how environmental conditions affect cultural group selection mechanisms. Cultural group selection may lead to community adoption of sustainable norms, where solitary groups fail to adopt cooperative behaviors. With this model, we can test under what environmental conditions this is true and what the implications are for community wellbeing and environmental protection.

## Expected Products
I plan to communicate the findings of this project both to the academic community and to non-scientists. The insights gained from this model will be published in a peer reviewed, academic journal such as Environmental Modeling and Software. This paper will show logically how we expect that protected areas function in the context of community level cooperative conservation. I also expect that this research will provide actionable insights for how designing protected areas while considering inter-group dynamics may allow for optimal resource provisioning under a much more broad set of circumstances than considering solitary groups alone.

To make this model readily available to a general audience however, I will update the online application provided above to better reflect the final version of the model. In this application I will also embed a graphical display of the model function as a .gif that displays key outputs at each timestep of the model. This will allow users to better internalize the implications of the model and to better apply them to real world conservation.

\newpage

# Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.

## Background
Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection [@molleman_cultural_2013]. One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use [@waring_coevolution_2017]. Under this prediction, selection acts both on individuals and on groups of individuals. When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection. Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups [@boyd_transmission_2010; @henrich_cultural_2004].

While these predictions are theoretically sound, they must be tested empirically [@kline2018designing]. Two such studies have begun to test the basis of cultural group selection in the natural world. The first demonstrated that culture offers greater fodder for group selection than do genetics. Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals. The proportion of variance in an allele or trait found between groups is referred to as genetic $F_{ST}$ or cultural $F_{ST}$ respectively. Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].

A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups. Researchers created pair-wise estimates of cultural $F_{ST}$ between 12 ethnic groups. These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups. Results of this study showed a strong association between cultural similarity and willingness to cooperate. This research indicates that the presence of variable cultural groups influences within-group cooperative behavior [@handley_human_2020].

This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly. Insights from cultural group selection may provide solutions to pressing environmental challenges of collective action. If validated experimentally, principles of cultural group selection may allow conservation scientists to better predict the viability of conservation initiatives.

Protected areas are an ideal study system for testing the mechanisms of cultural group selection in real world conservation. When a community adopts a protected area, like the Community Forest Management Agreements observed on Pemba Island, Tanzania, individuals agree to forego a portion of their personal harvest potential for the benefit of the community. Under cultural group selection, we only expect this behavior to emerge when the proportion of dissimilar to similar cultural traits between a group and its neighbors outweighs the benefit/cost ratio of engaging in cooperative conservation. Plainly put, communities should be more likely to adopt CoFMAs if they are culturally dissimilar to their surrounding communities. This project will test this theory by evaluating the perceived individual costs and group benefits of CoFMAs, within and between group cultural similarities (cultural $F_{st}$), and actual adoption of CoFMAs in forest dependent wards on Pemba Island [@kline2018designing].

### Question: 
Do Pemban wards with stronger cultural differentiation from surrounding wards adopt CoFMAs at a higher rate than wards who are more culturally similar to their neighbors?

### Predictions
* P1: Cultural differentiation of Pemban wards will predict CoFMA adoption only when the perceived individual costs of CoFMAs are relatively low.

* P2: High levels of forest dependence and perceived costs of CoFMAs will prevent CoFMA adoption, even when groups are considerably culturally dissimilar to adjacent groups.

## Research Design
Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level. These group level selection pressures are what we hypothesize to drive the adoption of CoFMAs. Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred. For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.

I will conduct an in-person survey of randomly sampled forest users in each Pemban ward. This survey will elicit information on approximately 50 cultural norms ranging from crime, cultural markers, punishment, family dynamics, etc. Cultural $F_{ST}$ values will be calculated for each trait given the following formula:

$$F_{STx} = \frac{\frac{n_{i}}{n_{i}+n_{j}} (p_{i}-\bar{p})^2 + \frac{n_{j}}{n_{i}+n_{j}} (p_{j}-\bar{p})^2}{\bar{p}(1-\bar{p})}$$

Where $i$ and $j$ are the two populations being compared, $n_{i}$ and $n_{j}$ represent the number of survey respondents for each group respectively. For trait $x$, $\bar{p}$ is the total frequency of trait $x$ in both populations, $p_{i}$ and $p_{j}$ are the trait frequency in each population. The equation can be interpreted as the variance in the relative frequency of trait $x$ between populations divided by the total variance in the trait across both populations. These values will be averaged across all traits to produce the total cultural $F_{ST}$ values between each pair of wards.

In addition to measuring the between and within group cultural variance, the in-person survey will also elicit information regarding individual’s perceived costs/benefits of CoFMAs, as well as group costs/benefits of CoFMAs. We can then use the cultural $F_{ST}$ values, paired with perceived costs and benefits to assess whether or not group cooperation will be favored by selection using the Rogers model of group selection by selective migration, from hereon referred to as the Rogers Inequality.

The Rogers Inequality model was initially designed to measure genetic evolution of groups via selective migration, but is equally applicable to, and has been adapted for cultural evolution [@bowles2006group]. This model presents an inequality under which individually costly, yet group beneficial traits are favored by selection. The emergence of group cooperative norms will be favored by selection when the ratio of between-group variance ($F_{ST}$) to within-group variance ($1-F_{ST}$) exceeds the ratio of individual costs of a behavior ($C_{w}$) to group level population benefit ($B_{p}$). Thus, if the following inequality is true, then individually costly, group beneficial behavior is favored by selection [@kline2018designing].

$$\frac{c_{w}}{b_{p}} < \frac{F_{ST}}{(1-F_{ST})}$$

While the primary output of this model is a true/false estimate of whether the emergence of a costly cooperative norm is favored by selection, we can also use the ratio presented in the inequality as a measure of selection strength. For example, we will expect Community Forest Management Agreements to be more strongly selected for when the right-hand side of the inequality is far greater than the left-hand side compared to when the sides are close to equal.

## Preliminary Work
This project benefits from preliminary work done by our collaborators at the Max Planck Institute for Evolutionary Anthropology (MPI). MPI researchers conducted a preliminary qualitative survey and collation of oral histories with the assistance of Department of Forests and Non-Renewable Natural Resources staff to confirm anecdotally that competition between wards has led to CoFMA adoption at the ward level. These researchers have also visited wards across the island and recorded which wards have adopted CoFMAs and which have not. These adoption characteristics will be compared against the adoption characteristics recorded in my work as a form of data validation.

## Expected Outputs
The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward. I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of CoFMAs. Figure \@ref(fig:selPress) shows an example of what the results of this study might look like. The findings of this study will be submitted for publication in a peer reviewed, academic journal.

```{r selPress, echo=FALSE, fig.cap="Simulated data showing protected area adoption vs calculated selection pressure for group adoption of protected areas.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/SelPressureAdopt.png")
```
\newpage

# Chapter 3: Social influences on transmission fidelity in household level diffusion of beneficial technology.

## Background
Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined [@muller2018state]. Household fuel production and global carbon sequestration are often seen as conflicting forest benefits. Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels [@mcshane_hard_2011]. Given limited funding, these projects are not intended to solve problems in their entirety, but rather to seed innovations that catalyze broad change. Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics [@keppel_conservation_2012; @oates_myth_1999; @stringer_land_2007].

This project will examine the spread of fuel-efficient cookstoves on Pemba Island, Tanzania (see Figure \@ref(fig:stovePic)). An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 [@lung_large-scale_2019; @sakthivadivel2018experimental]. The improved design has been subsequently copied throughout the communities and is thought to reduce deforestation and the time individuals — primarily women and children — must spend harvesting fuelwood.

```{r stovePic, echo=FALSE, fig.cap="Example of an improved, fuel efficient cookstove on Pemba Island, Tanzania. Image from Same Mwanga Environmental Conservation Advisory Organization", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/cookstove.jpg")
```

While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally. Although the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially [@lyons_hidden_2007; @mcguigan_over-imitation_2011]. In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly [@bartlett1932remembering]. Behavioral reconstruction creates variation in copied behaviors [@horner_faithful_2006]. Replication variation is thought to be greatly influenced by environmental factors, including social context [@mesoudi_multiple_2008].

Both theoretical and experimental studies have examined the relationship between transmission fidelity and environmental influences. Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning [@boyd_modeling_2002; @nakahashi_adaptive_2012]. This theory has also been validated experimentally [@muthukrishna_when_2016]. When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model [@king_when_2007].

While transmission fidelity can vary significantly given the context, it is often unclear what effect this variation has on the outcomes of imitated behavior. Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes [@lewis_transmission_2012]. Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes [@enquist_one_2010]. In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.

Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity. Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008]. This research will directly measure the impact of geographic and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.

In summary, conserving forest cover is critical for the global carbon balance and for the well-being of many rural communities. Conservation projects often fail due to a lack of consideration of the social processes underlying adoption. With few studies that measure transmission fidelity in a naturalistic setting, where outcomes have real economic and health impacts - we have an opportunity to explore how environmental constraints and social learning interact when outcomes matter. This project seeks to further investigate how and why behaviors are modified during the transmission process and how modification (both erroneous and intentional) affect performance targets.

### Question
What effect do geographic distance and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?

### Predictions
* P1: Transmission fidelity will increase with increased access to behavioral demonstrators through geographic proximity and/or social connectedness.

* P2: As access to behavioral demonstration decreases, the replication fidelity of non-functional stove traits will decrease strongly relative to functional traits.

## Research Design

### Approach
This project will take advantage of an ongoing initiative to spread fuel-efficient cookstoves on Pemba Island, Tanzania. This initiative has funded the building of a small number of improved cookstoves across communities which have subsequently been voluntarily copied by other community members. To answer the research questions above, I will collect data on the visual similarity of all copied cookstoves and the models they were copied from to assess direct transmission fidelity. I will also measure the fuel efficiency of the cookstoves as the behavioral outcomes of interest.

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

```{r mapp, echo=FALSE, fig.cap="Satellite map of Pemba Island, Tanzania. Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "75%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/InsetMap.png")
```

The survey instrument used for this research will elicit information regarding the relationship between the builder of each cookstove and the individual they copied their stove design from. Specifically, this survey tool will be used to gather data on geographic distance and communication frequency. Geographic distance will be measured according to the shortest walking distance between the two stoves. Communication frequency will be measured in terms of weekly communication and taken as a measure of social connectedness between the original stove builder and the copying individual. I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.

To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions]. Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test. Fuel efficiency will be calculated using the following thermal efficiency equation:

$$n_{th} = \frac{C_{p}*m_{wi}*(T_{f}-T_{i})+H_{v}*(m_{wi}-m_{wf})}{(F_{i}-F_{f})(1-\frac{M}{100})*LHV_{wood}-(F_{i}-F_{f})*\frac{M}{100}*(C_{p}*(T_{b}-T_{a})+H_{v})-LHV_{char}*C_{c}} $$

where $C_{p}$ represents the heat capacity of water (4.186 J/g-K), $m_{wi}$ is the initial mass of water used, $m_{wf}$ is the mass or remaining water, $T_{i}$ and $T_{f}$ are the initial and final water temperatures respectively, and $H_{v}$ represents the enthalpy of vaporization of water (2260 J/g). $F_{i}$ and $F_{f}$ represent the initial and final masses of fuel used. $M$ is the percent moisture content of the fuelwood used, $LHV_{wood}$ and $LHV_{char}$ are the lower heating values of wood and charcoal respectively, $T_{a}$ is the ambient temperature, $T_{b}$ is the local boiling temperature of water, and $C_{c}$ is the mass of the charcoal created during the test [@defoort2009stove].

While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics. I will categorize these characteristics as functional (e.g. stove size) or nonfunctional (e.g. decorative qualities). I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from in both functional and nonfunctional characteristics.

The key research questions for this project will be assessed using multilevel regression operationalized in a Bayesian context. I will compare standardized parameter estimates, to determine which factors (geographic distance or social relatedness) most strongly contribute to copying fidelity in this study system. Similar analyses will be used to determine which factors, if any, are strongly associated with stove fuel efficiency.

This project will contribute to scientific reproducibility in multiple ways. First, this work will be preregistered with the Open Science Framework. To confirm the analytical plan in advance of preregistration, I will fit the statistical models proposed above to simulated data and verify that the models recover the parameters known to have been generated in the simulated data. Secondarily, once all data for this project has been collected and the analyses are complete, all non-confidential data and R code used for this project will be made publically available through a GitHub repository shared by all team members.

## Preliminary Work
In preparation for a cancelled field season in the Summer of 2020, I created a suite of forms and documents to facilitate data collection. These include:

* A household livelihood and stove adoption survey.

* A datasheet for cookstove fuel efficiency measurements.

* A fuel efficiency measurement protocol.

I also simulated data that matches the data would be collected by the multiple instruments. These data were used to test the ability of my proposed statistical methods to recover the signal of demonstrator access on behavioral transmission fidelity and behavioral outcomes.

## Expected Outputs
The regression outcomes from this analysis will be the primary result presented in the publication of this research. Figures \@ref(fig:csp1) and \@ref(fig:csp2) show significant and insignificant theoretical outcomes that we might observe. The findings from this paper will also be shared broadly as a white paper describing how the insights from this research can help shape future conservation and development initiatives which distribute fuel efficient cookstoves in rural areas.

```{r csp1, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and weekly communication between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/CSP1.png")
```

```{r csp2, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and geographic distance between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch3/CSP2.png")
```
\newpage

# Conclusion
This research seeks to make clear how and why conservation behaviors succeed or fail to spread efficiently within a population. Moreover, the aim of this dissertation is to provide generalizable insights which can be applied to conservation initiatives in a variety of environments and population structures. A key advance made by this work is to link the fields of conservation social science and cultural group selection. Cultural group selection theory makes important predictions about how cooperative behaviors like conservation adoption may emerge and spread within populations, but has not yet been tested in a conservation context explicitly. Similarly, conservation social science incorporates environmental phenomena like resource mobility and regeneration, but largely omits generalizable insights about human behavioral evolution. By linking these two fields, we are greatly increasing the applicability of both bodies of work.

This dissertation will achieve this goal in three progressive steps. The first will formalize theory around cultural group selection in conservation contexts. This step is critical as conservation is a fundamentally different system from where these theories were originally developed. The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania. Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities to examine how behavioral efficiency may be altered by social contexts (e.g. communication frequency) as a behavior spreads throughout a population. All together, this research will allow us to better understand and possibly predict the emergence and effective spread of conservation norms within a population.

Human wellbeing is definitively tied to the persistence of renewable resources and ecosystem services. We are however, degrading these resources more rapidly than they can regenerate; reducing the stock of available resources while notably also slowing further regeneration. Conservation scientists have identified focal areas which, if preserved, may provide a disproportionate benefit to ecosystem function and offer a lifeline for human wellbeing. Simply put, we understand the environmental targets that we must achieve to ensure continued availability of renewable resources and we understand how conservation can achieve those targets. What we do not understand is how to effectively implement these required conservation measures. Reaching global biodiversity goals such as those set by the UN Convention on Biological Diversity will require “big and bold” conservation initiatives to be adopted at every level of the human population [@convention2010strategic; @dudley2018essential]. Currently however, even small-scale, local conservation measures often fail to gain traction and effectively run aground. This research will provide crucial theory and empirical evidence for how we can better shape and implement effective conservation initiatives at a variety of scales and in a variety of environmental conditions.

\newpage

# References