### 2021-11-20 DOI_Review/Manuscript.rmd
Text edits through quantitative applications

---
title: Utility of Diffusion Models for Predicting and Diagnosing the Spread of Conservation Behaviors
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: Boise State University- Human Environment Systems
  name: Vicken Hillis

date: "`r format(Sys.time(), '%B %d, %Y')`"
keywords: Community-based conservation, diffusion of innovations, cultural evolution, complex systems
bibliography: [./DOI_Conservation_Review.bib]
abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities. Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption. Here, we elaborate on the need for better understanding the spread of conservation behaviors, and review diffusion theory and its past application to these problems. We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks. Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest. Throughout this review, situate diffusion theory in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation. Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention. Overally, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion in these systems.

fontsize: 12pt
link-citations: yes
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{bm}
- \usepackage{amsmath}

---

```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')

setwd("C:/Users/Matthewclark989/Documents/Pemba_Project/DOI_Review")
```

> "Diffusion study...is something like the use of radioactive tracers in studying the process of plant   growth: it helps illuminate the process."
>
> `r tufte::quote_footer('--- Everett M. Rogers, Diffusion of Innovations, 2010')`

It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002]. The sweeping behavioral changes required to meet these targets will thus have to originate from and be maintained by the communities living in and around threatened natural areas themselves [@amel_beyond_2017; @catalano_learning_2019]. Community-based conservation projects aim to do just this by incentivizing rural, and often very poor, communities to implement these types of changes with little to no external oversight [@western_natural_1994]. While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote long-term behavioral changes which spread from person to person and even community to community [@brooks_assessing_2013; @horwich_community_2007].

While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004]. A major question in implementing these projects is if a specific intervention will indeed catch on in a population and if so, how quickly [@radeloff_hot_2013]? Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019].

# Diffusion Theory
Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time. Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010]. When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level. The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters. Therefore, at the core of diffusion theory is the observation that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently [@henrich_search_2001].

When modifying their beliefs or behaviors, people do not imitate any other individual at random [@bandura_social_1977]. Instead, they preferentially adopt traits by employing a variety of learning biases such as success-biased learning where individuals who are perceived as more successful are more likely to be copied than unsuccessful individuals [@boyd_culture_1988; @laland_social_2004]. As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are unequally copied and their frequency within a population changes. While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016]. This information affects individuals’ behaviors, which in-turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991]. Thus, cultural evolution provides an intuitive way to scale from individual learning and decision making to population level patterns and even population level behavior-environment feedbacks.

Diffusion of novel beliefs and behaviors is a specific instance of cultural evolution, where the trait being transmitted is new to the population and outcompetes some previous cultural variant, such as in the classic example of hybrid corn seed adoption in Iowa farmers [@mesoudi_cultural_2016; @ryan_acceptance_1950]. We observe these patterns through diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950]. The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010 ].

In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010 ]. This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001]. These two curve shapes however are not completely discrete, as the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].

![Canonical "S" and "R" diffusion curves. "R" curves are generally indicative of a higher proportion of individuals employing independent learning. "S" curves are generally indicative of a higher proportion of individuals employing social learning. ](./SvsR2.png)

# Diffusion Theory in Community-Based Conservation

Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019]. High-level studies have begun identifying diffusion theory as potentially useful for diagnosing why some community-based conservation interventions spread and persist, while others fizzle out. For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters. The authors point to a sigmoidal diffusion pattern to explain the slow start of many initiatives and suggest that practitioners may actively spread information on interventions (as in Abernethy et al. [-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004]. Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from diffusion theory as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].

As these high-level studies suggest, qualitative insights from diffusion theory have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world. In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] show that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives. Similarly, in Fort Collins, CO, researchers demonstrate that both the speed and extent of diffusion of information on native plant gardening likely increase with after motivating communication in the population [@niemiec_efficacy-based_2021]. Romero-de-Diego et al. [-@romero-de-diego_drivers_2021] use expert interviews to identify that observability of benefits was a key driver of the rapid scaling-up of wildlife management areas in Mexico. Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1]. This finding mirrors those of studies which look at more general cultural adoption across island chains, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].

These studies are compelling in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation, but have been critiqued as qualitative suggestions may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020]. Below, we make a quantitative argument for using diffusion theory to understand the spread of community-based conservation interventions. We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and/or what they can do to promote its spread. The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data, and diagnose when in the lifetime of an intervention individuals may be dropping out.

# Quantitative applications

The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [@smaldino_coupled_2020]. This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral spread [@cavalli-sforza_cultural_1981; @strang_adding_1991]. Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [@famil-rohani_development_2019; @smaldino_coupled_2020]. These compartmental differential equations assign every member of a population to a category (S, I, or R), depending on the proportion of individuals in the other categories and parameter values for the natural history characteristics of a given disease [@tolles_modeling_2020]. In this instantiation, the natural history parameter $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.

\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} \nonumber
\end{align}

When the natural history characteristics of some pathogen is known, the SI, SIR, SIS, & SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013]. For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time. We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].

Below, we develop a minimal agent-based model to highlight our core dynamics of interest. In this abstraction, we simulate the spread in adoption of a conservation intervention from just a few initial individuals, to a stable proportion of the population. We use this simulation to produce realistic data on long-term adoption dynamics under tightly coupled behavior-environment feedbacks unique to community-based conservation interventions. Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using the SIRS model shown above. We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation.

# Agent-based simulation
We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population. We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under two social learning mechanisms. The first is unbiased copying, where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019]. The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of more successful agents [@baldini_two_2013]. This mechanism is formally demonstrated in equation 1, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean pay-off, $\bar{P}$, of both behaviors at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].

\begin{equation}
 \text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}}  
\end{equation}

Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011]. We establish a degraded landscape/seascape (hereafter referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource. In the first time step, we introduce a conservation intervention which delineates 20% of the landscape as a conservation area and is initially adopted by just 2% of the population. In each subsequent time step, agents enrolled in the conservation program receive a reward, and continue to harvest just from the remaining 80% of the unconserved landscape. Agents not enrolled in the intervention harvest from the entire landscape and forgo the reward. After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior in the next time step [@baldini_success-biased_2012]. The conservation and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area. As behavioral adoption under success-biased learning is dependent on harvest success, the success of each behavior is determined largely by the amount of available resources in each area, and the amount of resources in each area is a result of behavioral adoption, this learning strategy is contingent on behavior-environment feedbacks, unlike unbiased adoption which occurs independent of harvest success.

![Conceptual diagram showing process at each timestep in the agnt-based simulation. This simulation was used to produce data on the adoption of conservation behaviors as a function of social learning and environmental outcomes.](./ConceptDiagram.png)

A visual description of this minimal model is shown in figure 1, and the full ODD+ model description can be found in the supplemental material (S1). Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/). As this simulation is not the emphasis of this study, we show just two runs below. The starting conditions of these runs are identical except for the social learning mechanism employed by the population. As the unbiased learning mechanism is independent of the resource levels, the adoption curve follows the traditional “S” pattern and nears fixation in the long-term. Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unconserved landscape are reduced following increased harvest pressure there as a result of the conservation intervention. The full output of this latter simulation is shown below in figure 2.

![One run of the agent based simulation used to produce data for analysis with the process -based SIRS model of behavioral diffusion. Starting conditions for this simulation were: 100 individuals, total landscape, carrying capacity of 10000 resources, 20% starting resource level, individual payment amount of 9 resources for enrollment in the conservation intervention, 2% initial enrollment, success-biased learning strategy at a strength of 1.1, 30% probability of resource mobility when carrying capacity reached, a maximum harvest value of 15, and 30% resource regeneration rate per time step.](./SiccessBiasedFullRun.png)

# Predicting adoption of conservation behaviors

With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described above to the first 30% of observations. We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time. Here, "infected" individuals are those that are enrolled in the conservation intervention.

$$ \text{Y}_{t} \sim Binomial(\text{N}, \text{I}_{t}) $$
\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t}\nonumber
\end{align}

$$\beta,\gamma,\chi \sim Normal(0,0.5)$$

For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 3). As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values alone, especially in the early stages of an intervention, is a practical and straightforward application of diffusion theory to community-based conservation interventions. For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate whether they should focus more on increasing rates of uptake or retention going forward.

![Markov chains for the three parameters of interest from the SIRS model fit to simulated data. The agent-based simulation used to produce the data employed environmentally mediated success-biased learning. A figure showing the Markov chains for the unbiased learning simulation can be found in the supplemental material (S3).](./SuccessPars.png)

We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for a community-based conservation initiative. In figures 4 & 5, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step. Grey and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively. These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning. R and STAN code for these models can be found in the supplemental material (S4).

![SIRS model fit of agent-based simulation data under unbiased social learning shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)

![SIRS model fit of agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)

Figures 4 & 5 also demonstrate that different learning mechanisms, while both producing a sigmoidal adoption pattern, can result in different overall rates in adoption of an intervention, even under identical starting conditions. In the simulation of a population employing unbiased learning, the population equilibrates at nearly 90% adoption of the intervention behavior. For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term.

# Shortcomings for application of diffusion theory to community-based conservation

As demonstrated above, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes. This model, while accurately capturing the pattern of spread, is unable to provide insight as to the mechanism of spread or causality in the system. As shown above and in existing literature, the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [@smaldino_adoption_2017; @andrews_cultural_2018; @tam_measuring_2021]. Therefore, while we agree that diffusion theory can be productively applied to community-based conservation, we point to theories of social learning and cultural evolution as a better alternative for diagnosing and planning interventions. For example, Tam et al. [-@tam_measuring_2021] and Kline et al. [-@kline_designing_2018] give practical examples of how to measure social learning strategies in the context of conservation, and tools (game/survey instrument, code, etc.) with which to collect and analyze the data.

Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020]. Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021]. Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].

The practice of applying statistical frameworks for parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly in ecology [@hefley_when_2017; @laubmeier_ecological_2020]. This emerging methodology allows for the estimation of interesting and directly interpretable process parameters, not possible with linear regression. We believe this is an exciting new avenue for conservation science as the direct interpretability and applicability of process parameters lend themselves well to tight integration with applied research as is common in the field. Future research may be interested in extending the SIRS diffusion model to include a term which represents change in the environment over time and the effect of resource levels on adoption rates to more accurately reflect the social-ecological coupling. An example of such an extension is presented below, where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$. We also must include three additional parameters: $\phi$, $\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.

\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} - \phi \cdot \text{E}_{t} \\
&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} + \phi \cdot \text{E}_{t} \nonumber\\
&\frac{dE}{dt} = - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta \cdot \text{E}_{t}\nonumber
\end{align}

While process models may be preferable to simulation and linear regression approaches for predicting adoption outcomes of conservation interventions, they might not always be possible. In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation and may even become mathematically intractable [@laubmeier_ecological_2020]. Further, as has been demonstrated in coastal fisheries, these systems can exhibit mathematical chaos, where our inability to measure the parameters of interest at a fine enough resolution results in seemingly stochastic, unpredictable outcomes [@wilson_chaos_1994].

Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007]. In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS or SIRS extension models presented above [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].

# Conclusions

Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply diffusion theory to community-based conservation. We show that these models can be powerful for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption. We also introduce cultural evolutionary research as a better set of theories for diagnosing the spread, or lack thereof, in such interventions and for planning future interventions. Lastly, we describe conservation as a complex adaptive system and propose simulation modeling as an alternative to process-based models when they may be analytically intractable or ill suited for system dynamics.

\newpage

# References