### 2021-12-19 DOI_Review/Manuscript.rmd
File path change 4 new comp

---
title: Utility of Diffusion of Innovations for Diagnosing the Spread of Conservation Behaviors
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: Boise State University- Human Environment Systems
  name: Vicken Hillis

date: "`r format(Sys.time(), '%B %d, %Y')`"
keywords: Community-based conservation, diffusion of innovations, cultural evolution, complex systems
bibliography: [./DOI_Conservation_Review.bib]
abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities. Diffusion of innovations theory (DOI) provides generalized heuristics and models that represent how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption. Here, we elaborate on the need for better understanding conservation institutions as a product of individual behavioral adoption and review DOI and its past application to these problems. We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks. Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest. Throughout this review, we situate DOI in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation. Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention. Overall, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion of community-based conservation initiatives.

fontsize: 12pt
link-citations: yes
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{bm}
- \usepackage{amsmath}

---

```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')

setwd("C:/Users/Matt/Documents/Pemba_Project/DOI_Review")
```

> "Diffusion study...is something like the use of radioactive tracers in studying the process of plant   growth: it helps illuminate the process."
>
> `r tufte::quote_footer('--- Everett M. Rogers, Diffusion of Innovations, 2010')`

It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002]. We are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas [@amel_beyond_2017; @catalano_learning_2019]. Community-based conservation projects aim to incentivize these behaviors in rural, and often very poor, communities with little to no external oversight [@western_natural_1994]. While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by communities who harvest natural resources to meet their daily needs [@brooks_assessing_2013; @horwich_community_2007]. While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions and manage environments for multiple interests is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999; @yates_purpose_2019].

Broadly defined, institutions, like those introduced by community-based conservation interventions, are socially constructed and transmitted prescriptions on behavior, which generally aim to incentivize collective action [@lott_review_1992; @currie_evolution_2016; @richerson_tribal_2009]. In natural resource management, the desired outcome of any particular institution varies on a case-by-case basis but may include goals like increasing the livelihood benefits produced from a resource or improving the biodiversity of a particular area [@agrawal_studying_2014]. Regardless of the exact social or ecological outcome of interest, “success” of a conservation institution has become synonymous with the degree to which the particular behavioral prescription is adopted by a population and whether or not that behavior is sustained over time [@agrawal_common_2001; @yates_purpose_2019].

Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in a population, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990]. These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, political structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009]. Although at least 40 unique factors have been identified as contributing to the success of community-based conservation institutions, well defined social and ecological boundaries, and well connected social systems stand out as major important themes [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001]. Research on common property additionally recommends that institutions be matched in scale to the underlying ecological processes of the resources they aim to manage [@mckean_common_2000].

Many of the factors that support the long-term viability of conservation institutions are intertwined such that the applicability of one may depend on the level of another and some factors may be inextricably correlated [@agrawal_group_2001; @agrawal_sustainable_2003]. Further, which specific factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for successful conservation institutions [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001]. Thus, even with decades of productive research on sustainable resource management, complete quantitative syntheses and insights have remained elusive.

We argue here that considering the stability of institutions within a population from the perspective of individuals, rather than entire communities may provide valuable insight for implementing community-based conservation projects, both qualitatively and quantitatively. Sustainable resource management institutions are observed at the community level, but in reality they are an emergent property of individual level decision making — communities do not adopt behaviors, individuals do [@currie_cultural_2021]. The normative conditions imposed on individuals by their ecological and social environments (i.e. the factors referenced above) affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention.

A major question in implementing these projects is if a specific intervention will indeed be adopted by members of a population and if so, how quickly [@radeloff_hot_2013]? Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019]. By examining sustainable resource management institutions as the emergent outcome of individual behavioral changes, we can leverage insights from other areas of social science to further guide theory on why some institutions emerge and others fail to be adopted long-term [@berkes_rethinking_2004].

In this article, we examine one social science theory specifically focused on individual behavior adoption, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010]. Our primary aim is to give a practical demonstration of how insights from DOI can be applied to expand our inference about the spread and persistence of conservation interventions beyond narrative descriptions. We do this by first simulating data on conservation adoption in a resource-dependent population and then statistically fitting a model of diffusion to these data. Adoption of resource safeguarding behavior is fundamentally different from behavioral adoption in the abstract in that individual adoptions affect the natural environment and alter the payoff structure under which individuals continue to adopt in future time periods. Our simulation highlights this feedback of interest in order to answer the key question of whether DOI is applicable to community-based conservation specifically, given this unique adoption dynamic. Throughout this article, we situate DOI theory in the broader literature on cultural evolution and argue that this body of work may provide more actionable insight than DOI in isolation, particularly by considering underlying mechanisms of behavioral transmission.

# Diffusion of innovations
Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time. DOI recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010]. When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level. The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters. Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior, rather than weighing the pros and cons independently [@henrich_search_2001].

When modifying their beliefs or behaviors, people do not imitate other individuals at random [@bandura_social_1977]. Instead, they preferentially adopt traits by employing a variety of learning biases, such as success-biased learning, where individuals who are perceived as more successful are disproportionately likely to be copied [@boyd_culture_1988; @laland_social_2004]. As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are unequally copied and their frequency within a population changes. While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016]. This information affects individuals’ behaviors, which in turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003]. Thus, cultural evolution provides an intuitive way to integrate  individual decision-making, population level patterns, and even population level behavior-environment feedbacks.

Diffusion of novel beliefs and behaviors is a specific instance of cultural evolution, where the trait being transmitted is new to the population and outcompetes some previous cultural variant, such as in the classic example of hybrid corn seed adoption in Iowa farmers [@mesoudi_cultural_2016; @ryan_acceptance_1950]. We observe these patterns using diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950]. The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].

In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010]. This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001]. These two curve shapes however are not completely distinct. As the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].

![Canonical "S" and "R" diffusion curves. "R" curves are generally indicative of a higher proportion of individuals employing individual learning. "S" curves are generally indicative of a higher proportion of individuals employing social learning. ](./SvsR2.png)

# Diffusion of innovations in community-based conservation
Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019]. Fittingly, high-level studies have begun identifying DOI as potentially useful for diagnosing why some community-based conservation interventions spread and persist, while others fizzle out. For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters at the sigmoidal equilibrium. The authors take insights from DOI to suggest that practitioners may actively spread information on interventions (as in Abernethy et al. [-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004]. Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from DOI as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].

As these high-level studies suggest, qualitative insights from DOI have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world. In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] showed that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives. Similarly, in Fort Collins, CO, researchers demonstrated that both the speed and extent of diffusion of information on native plant gardening likely increase when community members are offered incentives to increase communication [@niemiec_efficacy-based_2021]. Romero-de-Diego et al. [-@romero-de-diego_drivers_2021] used expert interviews to identify that observability of benefits was a key driver of the rapid adoption of a voluntary wildlife management program among landholders in Mexico. Conversely, in contexts where observability and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1]. This finding mirrors those of studies which look at more general cultural adoption across island chains, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].

These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI that may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020]. Below, we make a quantitative argument for using DOI to understand the spread of community-based conservation interventions. We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and what they can do to promote its spread. In this review, we therefore explicitly demonstrate how DOI can be used to generate predictions about the patterns of uptake of conservation behaviors given behavior-environment feedbacks and diagnose how an intervention may be best improved.

# Quantitative applications
In this section we explain how we can fit an observed pattern of diffusion mathematically to estimate the rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population. The “S” shaped diffusion pattern of behavioral spread over time mirrors that of community disease transmission [@smaldino_coupled_2020]. This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral spread [@cavalli-sforza_cultural_1981; @strang_adding_1991]. Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [@famil-rohani_development_2019; @smaldino_coupled_2020]. These compartmental differential equations assign every member of a population to a category (S, I, or R), depending on the proportion of individuals in the other categories and parameter values for the natural history characteristics of a given disease [@tolles_modeling_2020]. In this instantiation, the natural history parameter $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.

\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
&\frac{dR}{dt} = \gamma\cdot \text{I}_{t} - \chi \cdot \text{R}_{t}\nonumber
\end{align}

When the natural history characteristics of some pathogen are known, the SI, SIR, SIS, & SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013]. For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time. Instead, the qualitative insights from DOI and the study of sustainable resource management institutions referenced above aim to increase the rate of spread and reduce the rate of dropout so that the intervention persists in the population and may even spread to new communities. While we can not know exactly what the rates of spread and dropout will be ahead of time, we can use Bayesian inference to probabilistically estimate them once the project has been implemented [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018]. Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are limiting the overall rate of uptake, and apply the qualitative insights from DOI and natural resource management institutions in a more targeted fashion.

The uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption generally [@liu_complexity_2007]. In the case of marine protected areas for example, success in conserving fish populations may actually disincentivize individuals to continue supporting conservation programs [@christie_marine_2004]. A key question then is whether the diffusion models explained above will still reliably fit the spread of conservation behaviors when those behaviors alter the environment, and changes in the environment alter the payoff structure of various behavioral choices. To answer this question, we use an agent-based simulation to generate data on the spread of a conservation behavior via social learning, contingent on feedbacks with the environment. Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using a Bayesian instantiation of the SIRS model shown in equation 1. We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation. While the purpose of this example is to test whether the SIRS model can capture our core dynamics of interest, it also serves as an explicit demonstration of how a practitioner or other researchers can directly apply the quantitative process described above. We first briefly describe the simulation below.

# Agent-based simulation
We seek to test whether the SIRS model shown in equation 1 can produce reliable insights about community uptake of a conservation behavior, especially given behavior-environment feedbacks common in natural resource management institutions. To answer this question, we first need complete time series data in-hand which are a product of our dynamics of interest. To obtain these data, we develop a minimal agent-based model which we use to simulate data on the spread of a resource safeguarding behavior within a population. For comparison, we simulate adoption data both independent of, and contingent on feedbacks with the environment. These two dynamics are operationalized using two different social learning mechanisms for behavioral transmission [@kendal_social_2018]. The first is frequency-dependent copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the harvest benefit provided by adopting the behavior or not [@barrett_equifinality_2019]. We include a simulation using this mechanism as a baseline for how behavior-environment feedbacks inherent to our second mechanism affect the overall pattern of adoption. The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of other agents with the greatest harvesting success [@baldini_two_2013]. This mechanism is formally demonstrated in equation 2, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean harvest, $\bar{P}$, from each behavior at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019]. Under biased social learning, the strength of the learning biases (e.g. $\beta$) control how disproportionately agents copy others based on a given condition (e.g. success); a $\beta$ = 1 in this formulation would thus mean that the probability that one agent copies another is directly proportional to the ratio of their harvests.

\begin{align}
\text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}} 
\end{align}

Our simulated system is conceptualized as a fishery where agents are directly compensated for forgoing resource extraction from a particular area if they choose [@rakotonarivo_role_2021]. We establish a fishery, loosely based on a global average stock integrity, at 20% of its potential carrying capacity [@fao_state_2020]. In the first time step, we introduce an intervention which delineates 20% of the fishery as a conservation area and just 2% of agents initially change their behavior to forgo harvesting there. Fish reproduce proportionally to the population in each time period, and move freely between the conservation area and unconserved area. In each subsequent time step, agents who adopted the resource safeguarding behavior (enrolled) receive a reward, and continue to harvest just from the remaining 80% of the unconserved fishery. Agents not enrolled in the intervention harvest from the entire fishery and forgo the reward. After each harvest period, agents are randomly paired, and employ either frequency-dependent or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012]. The conserved and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area. A visual description of this minimal model is shown in figure 2 and interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).

![Conceptual diagram showing process at each time step in the agnt-based simulation. This simulation was used to produce data on the adoption of conservation behaviors as a function of social learning and environmental outcomes.](./ConceptDiagram.png)

We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-dependent and success biased. Aside from the learning mechanism employed by the population, all model conditions are identical for these two scenarios. Frequency-dependent transmission is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment. Under the success-biased learning scenario, agents compare their harvests to others in each time step and disproportionately copy more successful agents. In this instantiation then, adoption is dependent on harvest success and harvest success is determined largely by the amount of available resources in each area. As the amount of available resources in an area is determined by the number of agents who harvest there, this learning mechanism incorporates behavior-environment feedbacks typical of natural resource management institutions. Because agents compare their own harvests to those of others in each time step under the success-biased learning scenario, rather than exclusively imitating other individuals, this scenario actually represents a mix of individual and social learning. The adoption pattern observed in this simulation results in a mix between an “R” and “S” type curve, reinforcing the interpretation of this scenario as a mix of individual and social learning [@henrich_cultural_2001]. Figure 3 shows the full output of the later simulation where adoption (enrollment) spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.

![One run of the agent-based simulation used to produce data for analysis with the SIRS model of behavioral adoption under the success-biased learning scenario. The top figure shows the number of resources in both the conservation and unconserved areas over time, as a percent of the carrying capacity in those areas. The middle figure shows the average number of resources harvested (payoff) by each agent at each time step. The bottom figure shows the number of agents who have adopted the conservation behavior (enrolled) at every time step.](./SiccessBiasedFullRun.png)

# Predicting adoption of conservation behaviors
With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described in equation 1 to the first 30% of observations. We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation (3) below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time. Here, "infected" individuals are those that are enrolled in the conservation intervention.

$$ \text{Y}_{t} \sim Binomial(\text{N}, \text{I}_{t}) $$
\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t}\nonumber
\end{align}

$$\beta,\gamma,\chi \sim Normal(0,0.5)$$

For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 4). As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values, especially in the early stages of an intervention, provides practical and straightforward intuition to diagnose areas for improvement in community-based conservation interventions. For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate whether they should focus more on increasing rates of uptake or retention going forward. In addition, this insight may help practitioners better decide which factors from the institutional design literature referenced above may be most impactful for their specific system.

![Markov chains for each of the three parameters of interest from the SIRS model fit to simulated data. Beta is interpretable as the rate of adoption of the behavior (enrollment). Gamma is an estimate of the rate of dropout. Chi represents the rate at which individuals who previously dropped out are added back into the susceptible category. The agent-based simulation used to produce the data employed environmentally mediated, success-biased learning. A figure showing the Markov chains for the unbiased learning simulation can be found in the supplemental material (S3).](./SuccessPars.png)

We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for a community-based conservation initiative. In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step. Gray and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively. These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning. R and STAN code for these models can be found in the supplemental material (S3).

![SIRS model fit of agent-based simulation data under unbiased social learning shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)

![SIRS model fit of agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)

Figures 5 & 6 also demonstrate that different scenarios and respective learning mechanisms, while both producing fixation in the adoption rate, can result in different overall rates in adoption of an intervention, even under identical starting conditions. In the simulation of a population employing frequency-dependent learning, the population equilibrates at nearly 90% adoption of the intervention behavior. For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term. Lastly, the mix of social and independent learning employed by agents in the success-biased learning scenario produces a somewhat “R” shaped curve as explained in the section above. These results show that the SIRS model of diffusion can reliably fit the pattern of behavioral adoption even when the pattern does not follow a strict sigmoidal “S” shape.

# Shortcomings for application of diffusion of innovations to community-based conservation
As demonstrated in figures 5 & 6, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes. Fitting this model to observed data on the uptake of a behavior can provide us with estimates of the rates of adoption, dropout, and readoption for a specific intervention and allow us to predict the final stable rate of adoption in the population from early in the project. These insights, while useful, are somewhat limited in their ability to prescribe improvements for an intervention. As highlighted by the quote in the beginning of this article, diffusion study can explain what is happening in a system, but gives little intuition as to the mechanisms behind why it is happening [@rogers_diffusion_2010].

As first described by Henrich [-@henrich_cultural_2001] and demonstrated in figures 5 & 6, we see clear changes in the population patterns of adoption when individuals employ varying levels of social or independent learning. But when comparing many biased social learning mechanisms, the pattern of adoption may exhibit equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019]. Critically, this makes it virtually impossible to infer the specific biased social learning mechanism(s) used by a population from the pattern of adoption over time alone. This information is however important for researchers and practitioners, as the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [@smaldino_adoption_2017; @andrews_cultural_2018; @tam_measuring_2021]. Plainly, DOI can tell us if an intervention is being robustly adopted or not, but we need to understand the actual transmission mechanism to effectively act on this information. Thankfully, this information is possible to collect. For example, Tam et al. [-@tam_measuring_2021] and Kline et al. [-@kline_designing_2018] give practical examples of how to measure social learning strategies in the context of conservation, and tools (game/survey instrument, code, etc.) with which to collect and analyze the data.

With information on the patterns and mechanisms of uptake of community-based conservation behaviors we can more selectively apply the many narrative insights from DOI and the study of natural resource management institutions. In populations primarily employing prestige-biased learning (where prestigious individuals are disproportionately copied) for instance, the identification of and collaboration with community change agents may be particularly effective [@jimenez_prestige-biased_2019; @henrich_evolution_2001; @dietz_drama_2002]. Conversely, the observability of both resources and the benefits of conservation are likely of increased importance when the population primarily employs content-biased learning (where some characteristic of the trait itself makes it more desirable than a previous trait)  [@ostrom_governing_1990; @rogers_diffusion_2010; @boyd_culture_1988-1].

# Future work
Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020]. Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021; @le_polain_de_waroux_role_2021]. Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].

Mechanistic, or process models allow for us to formally describe these complex interactions while testing the logical consequences of our assumptions and the relation of our hypothesized processes to observed data [@epstein_why_2008]. The practice of applying statistical frameworks for parameter estimation to analytical process models, and in particular to ordinary differential equations, is however still relatively new and has been embraced mostly in ecology [@hefley_when_2017; @laubmeier_ecological_2020; @morales_mechanistic_2021]. This emerging methodology allows for the estimation of interesting and directly interpretable process parameters, not possible with linear regression. We believe this is an exciting new avenue for conservation science as the direct interpretability and applicability of process parameters lend themselves well to tight integration with applied research as is common in the field. Future research may be interested in extending the SIRS diffusion model to include a term which represents change in the environment over time and the effect of resource levels on adoption rates to more accurately reflect the social-ecological coupling. An example of such an extension is presented below, where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$. We also must include three additional parameters: $\phi$, $\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.

\begin{align}
&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} - \phi \cdot \text{E}_{t} \\
&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} + \phi \cdot \text{E}_{t} \nonumber\\
&\frac{dE}{dt} = - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta\cdot\text{E}_{t}\nonumber
\end{align}

While process models may be preferable to simulation and linear regression approaches for predicting adoption outcomes of conservation interventions, they might not always be possible. In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation and may even become mathematically intractable [@laubmeier_ecological_2020]. Further, as has been demonstrated in coastal fisheries, these systems can exhibit mathematical chaos, where our inability to measure the parameters of interest at a fine enough resolution results in seemingly stochastic, unpredictable outcomes [@wilson_chaos_1994].

Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007]. In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS  or SIRS extension model presented in equation 4 [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015]. Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; joffre_combining_2015]. As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of government are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].

# Conclusions
Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply DOI to community-based conservation. We test the utility of the SIRS model of diffusion on simulated data which highlight the feedbacks between behavioral adoption and environmental change commonly seen in community-based conservation interventions. We show that this model can be a powerful tool for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption, even under reciprocal  behavior-environment interactions. We also introduce the broader field of cultural evolution as a more comprehensive set of theories for understanding the underlying mechanisms of behavioral adoption than DOI alone. We hope that this review provides a model for how practitioners might practically apply insights from DOI to community-based conservation interventions, while also clarifying the position of DOI in the current literature on the adoption of sustainable resource management practices.

\newpage

# References